Analyzing Information Transfer from Different Data Sources via Precise Bias-variance Asymptotics

### 2021-07-08 revision_journal/exp.tex
finish sec 3.1

%\subsection{Comparison between the averaging and HPS estimators}
\subsection{Regression tasks}

\subsection{Text classification tasks}
Our results and simulations are all in the high-dimensional linear regression setting.
How well do they extend to other scenarios?
In this section, we conduct further studies on six text classification datasets.
Our datasets include a movie review sentiment dataset (MR) \cite{pang2005seeing}, a sentence subjectivity dataset (SUBJ) \cite{pang2004sentimental}, a customer reviews dataset (CR) \cite{hu2004mining}, a question type dataset (TREC) \cite{li2002learning}, an opinion polarity dataset (MPQA) \cite{wiebe2005annotating}, and the Stanford sentiment treebank (SST) dataset \cite{socher2013recursive}.
%The question is to predict positive or negative sentiment expressed in the text.
Our model consists of a word embedding layer with GloVe embeddings \cite{pennington2014glove} followed by a long-short term memory (LSTM) or a multi-layer perception (MLP) layer \cite{lei2018simple}.\footnote{For MLP, we apply an average pooling layer over word embeddings.
For LSTM, we add a shared feature representation layer on top of word embeddings.}

>>>>

\subsection{Evaluating HPS under model and covariate shifts}

\begin{figure}[!t]
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/compare_risk_covariate_shift.eps}
		\caption{Covariate shift}
		\label{fig_sec5_covariate}
	\end{subfigure}
	\caption{Comparing HPS to several natural estimators for transfer learning. We find that HPS enjoys superior performance under covariate and/or model shifts. The results are averaged over five random seeds.}
	\label{fig_sec51}
\end{figure}

We show that HPS estimators enjoy superior empirical performance compared to several natural transfer learning estimators.
We consider the following estimators for this comparative study:
\begin{itemize}
    \item {\bf Averaging estimator (AVG):} given two tasks, take a convex combination of their OLS estimators $b \cdot \hat{\beta}^{\STL}_1 + (1 - b) \cdot \hat{\beta}^{\STL}_2$.
    \item {\bf STL estimator (STL):} simply use the OLS estimator of task two $\hat{\beta}^{\STL}_2$ without using task one's data.
    \item {\bf Transfer learning estimator (TL):} simply use the OLS estimator of task one $\hat{\beta}^{\STL}_1$ without using task two's data.
\end{itemize}

Figure \ref{fig_sec51} shows the result under covariate and/or model shifts.
Figure \ref{fig_sec5_covariate} considers the same example as Figure \ref{fig_sec31}.
%    \begin{align}
%    L(\hat{\beta}^{\AV}(\lambda) ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \lambda^2 \left\| (\Sigma^{(2)})^{1/2}(\beta^{(1)}-\beta^{(2)})\right\|^2 \nonumber \\
%    &+  \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2   \lambda^2 \bigtr{\Sigma^{(2)} [(X^{(1)})^\top X^{(1)}]^{-1}  }\nonumber\\
%    & + \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2  (1-\lambda)^2 \bigtr{\Sigma^{(2)} [(X^{(2)})^\top X^{(2)}]^{-1}  }  . \label{L_AVE_simple}
%    \end{align}
%    and
%    \begin{align}
%    L(\hat{\beta}_2^{\STL} ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2  \bigtr{\Sigma^{(2)} [(X^{(2)})^\top X^{(2)}]^{-1}  }  . \label{L_STL_simple}
%    \end{align}
    %Moreover, there exists a constant $C_0>0$ such that with high probability,
    %\be\label{loss_large}
    %g(a) > g(0) \quad \text{for all $a$ such that }|a|\ge C_0.
    %\ee
    %with high probability for any small constant $\e>0$.

 %With Lemma \ref{fact_tr}, we obtain that w.h.p.,
%\begin{align}
%L(\hat{\beta}_2^{\STL} ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \frac{p \sigma^2 }{n_2-p}   , \label{L_STL_simple01}
%\end{align}
%and
%\begin{align}
%L(\hat{\beta}^{\AV}(\lambda) ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \lambda^2 \left\| (\Sigma^{(2)})^{1/2}(\beta^{(1)}-\beta^{(2)})\right\|^2 \nonumber \\
%&+  \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2 \left[\frac{ \lambda^2}{n_1-p}\bigtr{\Sigma^{(2)}(\Sigma^{(1)})^{-1}} + \frac{(1-\lambda)^2p  }{n_2-p}\right] . \label{L_AVE_simple01}
% \end{align}

\todo{merge into previous paragraphs {Simulations.}}
Finally, we validate the result of Example \ref{ex_covshift}.
Figure \ref{fig_covariate} shows task two's prediction loss as we increase task one's sample size.
Recall that $\lambda$ measures the severity of covariate shifts---a larger $\lambda$ means a larger covariate shift.
We indeed observe the dichotomy in Example \ref{ex_covshift} at $n_1 = 800$.
The sample size $n_2$ is $800$ and the noise variance $\sigma^2$ is $1/4$.

%{\cob The idea of proof: proved using free probability; can also use polynomials of random matrices.}

\subsection{Implications for regression methods}

\subsection{Implications for text classification}
Our results and simulations are all in the high-dimensional linear regression setting.
How well do they extend to other scenarios?
In this section, we conduct further studies on six text classification datasets.
Our datasets include a movie review sentiment dataset (MR) \cite{pang2005seeing}, a sentence subjectivity dataset (SUBJ) \cite{pang2004sentimental}, a customer reviews dataset (CR) \cite{hu2004mining}, a question type dataset (TREC) \cite{li2002learning}, an opinion polarity dataset (MPQA) \cite{wiebe2005annotating}, and the Stanford sentiment treebank (SST) dataset \cite{socher2013recursive}.
%The question is to predict positive or negative sentiment expressed in the text.
Our model consists of a word embedding layer with GloVe embeddings \cite{pennington2014glove} followed by a long-short term memory (LSTM) or a multi-layer perception (MLP) layer \cite{lei2018simple}.\footnote{For MLP, we apply an average pooling layer over word embeddings.
For LSTM, we add a shared feature representation layer on top of word embeddings.}

### 2021-07-08 revision_journal/intro.tex
finish sec 3.1

These previous works mainly consider cases where there is only one high-dimensional sample covariance matrix involved. %need to check 
In the single task case, if the data is Gaussian, then the precise variance limit can be derived directly from properties of Wishart distribution. In the high-dimensional setting, the eigenvalues of Wishart matrices satisfy the famous Marchenko–Pastur (MP) law \cite{MP}, whose Stieltjes transform characterizes the variance limit. Furthermore, it is well-known that the MP law holds universally regardless of the detailed data distribution \cite{bai2010spectral}. \cite{isotropic} obtained a sharp converge rate of the empirical spectral distribution (ESD) to the MP law for sample covariance matrices with isotropic covariance. \cite{Anisotropic,DY} later extended this result to sample covariance matrices with arbitrary population covariances. These results are proved by establishing certain optimal convergence estimates of the Stieltjes transforms of sample covariance matrices, which are referred to as \emph{local laws} in the random matrix theory literature. We refer the reader to \cite{erdos2017dynamical} and the references therein for a detailed review of related concepts. In this paper, we extend these techniques to the two-task setting and prove the local law for the sum of two sample covariance matrices with an arbitrary covariate shift, which is one technical contribution of this work. From the local law, we can derive the variance limit with a precise dependence on the singular values of the covariate shift matrix.

@@

On the other hand, the bias limit is much harder to derive. However, when the data is Gaussian distributed and there is no covariante shift, we can use recent results in ... to derive the bias limit. To the best of our knowledge, we are not aware of any previous results in the multi-task case, even for cases without covariate shifts.

>>>>

These previous works mainly consider cases with only one high-dimensional sample covariance matrix involved. %need to check 
In the single task case, if the data is Gaussian, then the precise variance limit can be derived directly from properties of Wishart distribution. In the high-dimensional setting, the eigenvalues of Wishart matrices satisfy the famous Marchenko–Pastur (MP) law \cite{MP}, whose Stieltjes transform characterizes the variance limit. Furthermore, it is well-known that the MP law holds universally regardless of the detailed data distribution \cite{bai2010spectral}. \cite{isotropic} obtained a sharp converge rate of the empirical spectral distribution (ESD) to the MP law for sample covariance matrices with isotropic covariance. \cite{Anisotropic,DY} later extended this result to sample covariance matrices with arbitrary population covariances. These results are proved by establishing certain optimal convergence estimates of the Stieltjes transforms of sample covariance matrices, which are referred to as \emph{local laws} in the random matrix theory literature. We refer the reader to \cite{erdos2017dynamical} and the references therein for a detailed review of related concepts. We extend these techniques to the two-task setting and prove a local law for the sum of two sample covariance matrices with an arbitrary covariate shift, which is one technical contribution of this work. From the local law, we can derive the variance limit with a precise dependence on the singular values of the covariate shift matrix.

@@

On the other hand, the bias limit is much harder to study, and its exact dependence on the sample ratio and the covariate shift matrix is still an open problem in random matrix theory in the most general setting. However, when the data is Gaussian distributed and there is no covariante shift, we can use recent results in \cite{BES_free1,BES_free2} to derive the bias limit. To the best of our knowledge, we are not aware of any previous result in the high-dimensional setting, even for two sample covariance matrices with isotropic population covariances and the same sample size.

%previous results in the multi-task case, even for cases without covariate shifts.
